---
title: "new_feature"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{new_feature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mGWASR)
library(stringr)
```

# Quick start

## get GWAS ID from ieu OpenGWAS database

get GWAS ID from OpenGWAS.

Please get OpenGWAS API token from <https://api.opengwas.io/>

```{r}
Sys.setenv(OPENGWAS_JWT="your API token")
```

```{r}
allGWAS <- ieugwasr::gwasinfo()
```

Alternatively, obtain the GWAS ID from the downloaded table in this R package.

```{r}
allGWAS_file <- system.file("extdata", "2024_allGWAS.csv", package = "mGWASR")
allGWAS <- read.csv(allGWAS_file,row.names = 1)
```

## Find causal SNPs through finemap

Select the chromosome regions of interest

```{r}
region <- '11:58780549-62223771'
```

Searching for causal SNPs of a specific GWAS in regions of interest

```{r include=FALSE}
causalsnp <- findcausalSNP(region,'ebi-a-GCST007515',threshold = 0.8, bfile = 'bfile/EUR') #Please specify the bfile file path
```

## Find co-localization relationships in large batches

Extract the GWAS ID of the metabolites

```{r}
pheno2 <- allGWAS[stringr::str_detect(allGWAS$id, "met-a|met-c"),'id']
```

Search for co localization relationships between pheno1 and multiple pheno2 in regions centered around causal SNPs (which can be multiple)

```{r}
batch_coloc('ebi-a-GCST007515', pheno2, rsid_or_pos = c('rs174537','rs174547','rs1535'),region = 500000,output_dir = 'output_dir') #region is usually 500kb-2Mb
```

Perform single co localization and plot

```{r}
single_res <- colocal('ebi-a-GCST007515',pheno2[1],rsid_or_pos = 'rs174537',region = 500000,plot = T)
#Alternatively, you can use a VCF file
#single_res <- colocal('pheno1_vcf_file','pheno2_vcf_file',rsid_or_pos = casusalsnp[1],region = 500000,plot = T)
```

## Perform LDSC analysis in batches conveniently to estimate heritability and genetic correlation

Download GWAS sumstats of interest

```{r}
phenos <- allGWAS %>%
  filter(grepl("Type 2 diabetes", trait, ignore.case = TRUE) & population == "European")
download_ieuvcf(phenos$id,'T2D_EUR_vcf/')
```

Filter successful downloads

```{r}
phenos <- download_status %>%
  filter(status == "Success")
```

Convert VCF files to intermediate RDS files required for LDSC analysis

```{r}
ldsc_file <- vcf2ldsc('T2D_EUR_vcf/',saveRDS = T,output_dir = 'T2D_EUR_vcf_rds',num_cores = 5)
```

Batch calculate the heritability of each phenotype

```{r}
ldsc_h2 <- batch_ldsc_h2('T2D_EUR_vcf_rds/',num_cores = 8)
```

Pair all objects pairwise and calculate the genetic correlation between every two

```{r}
ldsc_res <- pair_ldsc('T2D_EUR_vcf_rds/',output_dir = 'T2D_EUR_res/',para_plan  = 'multicore',num_cores = 8)
```

Merge genetic correlation results and filter them based on conditions to create a table for drawing

```{r}
com_res <- combine_filter('T2D_EUR_res/',traits = c('ebi-a-GCST005047.vcf',
  'ebi-a-GCST005413.vcf',
  'ieu-a-1090.vcf',
  'ukb-b-13806.vcf',
  'ebi-a-GCST007517.vcf'))
```

Plot a correlation heatmap

```{r}
plot <- cor_plot(com_res)
```

# Case study of T2D

This section is used to reproduce the results of the case study of T2D in the article, which may take a long time.

obtain the GWAS ID

```{r}
allGWAS_file <- system.file("extdata", "2024_allGWAS.csv", package = "mGWASR")
allGWAS <- read.csv(allGWAS_file,row.names = 1)
```

Find causal SNPs through finemap

```{r}
region <- c("11:58780549-62223771", "11:116383348-117747110", "19:44744108-46102697", 
  "11:59251804-62201641", "15:58441366-59694116", "1:54226262-56413117", 
  "11:116383543-117901740", "16:55903774-57664330", "2:26894985-28598777", 
  "8:19492840-20060856")
```

Searching for causal SNPs of a specific GWAS in regions of interest

```{r include=FALSE}
causalsnp <- findcausalSNP(region,'ebi-a-GCST007515',threshold = 0.8, bfile = 'bfile/EUR') #Please specify the bfile file path
#Preserve unique SNPs
causalsnp <- unlist(strsplit(na.omit(c(as.matrix(causalsnp))), ",\\s*"))
```

Find co-localization relationships in large batches

```{r}
pheno2 <- allGWAS[stringr::str_detect(allGWAS$id, "met-a|met-c"),'id']
```

Search for co localization relationships between pheno1 and multiple pheno2 in regions centered around causal SNPs (which can be multiple)

```{r}
batch_coloc('ebi-a-GCST007515', pheno2, rsid_or_pos = causalsnp,region = 500000,output_dir = 'output_dir') #region is usually 500kb-2Mb
```

Perform single co localization and plot

```{r}
single_res <- colocal('ebi-a-GCST007515','met-c-936',rsid_or_pos = 'rs769449',region = 500000,plot = T)
#Alternatively, you can use a VCF file
#single_res <- colocal('pheno1_vcf_file','pheno2_vcf_file',rsid_or_pos = casusalsnp[1],region = 500000,plot = T)
```

## 
